---
title: "Dead and Core reproduction"
author: "Jeffrey M. Young"
date: "15 May 2020"
output:
  pdf_document:
        latex_engine: xelatex
theme: prettydoc
---

In this walk-through we'll perform the plot generation for the dead and core
analysis. We assume you have read the other walk-throughs.

### Preliminaries
First, libraries, note that I have silenced the shadowing warnings from R:
```{r message=FALSE}
library(ggplot2)   #plotting
library(dplyr)     #dataframe manipulation
library(tidyr)     #dataframe manipulation
library(broom)     #for the tidy function
library(scales)    #for scientific function
library(latex2exp) #for TeX in Labels
library(ggpubr)    #for stat_cor
```

And now we'll load the data, and perform some simple munging to select relevant
columns.

```{r data}
## Load files
finResultsFile <- "../../data/fin_dead_core.csv"
autoResultsFile <- "../../data/auto_dead_core.csv"

## The usual manipulations
finData <- read.csv(file=finResultsFile) %>%
        mutate(Algorithm = as.factor(Algorithm), Config = as.factor(Config)) %>%
        mutate(Algorithm = gsub("-->", "\U27f6", Algorithm))

autoData <- read.csv(file=autoResultsFile) %>%
         mutate(Algorithm = as.factor(Algorithm), Config = as.factor(Config)) %>%
         mutate(Algorithm = gsub("-->", "\U27f6", Algorithm))

## combine the data frames
finDF <- finData %>% mutate(data = "Fin")
autoDF <- autoData %>% mutate(data = "Auto")

data <- rbind(finDF, autoDF) %>%
  select(Mean, Algorithm, data, ChcCount, PlainCount, Config)
```

Now let's look at the dataset

```{r peek}
str(data)

data
```

We can already see enough to reproduce the table in the paper, but some
formatting would be nicer:

```{r formatting}
deadCoreDF <- data %>%
  # perform the following actions for each dataset
  group_by(data) %>%
  # sort by mean
  arrange(Mean) %>%
  ## change the significant figures by dataset
  mutate(MeanLbl = case_when(data == "Auto" ~ signif(Mean, 5),
                             data == "Fin" ~ signif(Mean, 3)))

deadCoreDF
```

and a simple bar plot to show the data:

```{r deadCore, fig.width=7, fig.height=4, warning=FALSE}
ggplot(deadCoreDF , mapping = aes(x=Algorithm , y=Mean , color = Algorithm)) +
  theme_classic() +
  ## remove the legends for fill, size, color
  guides(fill = FALSE) +
  guides(size = FALSE) +
  guides(color = FALSE) +
  ## make the plot a bar plot
  geom_bar(stat="identity", fill="white", width=0.5
           ,size=1.15) +
  ## add the text labels to show exact values
  geom_text(aes(label=MeanLbl, hjust=1.4), color="black") +
  ## construct the plot for both data sets
  facet_wrap(. ~ data, scales = "free_x") +
  ## other niceties
  labs(title = "Dead Core Demonstration",
       y = "Time [s] to solve Dead and Core Analysis over all Versions") +
  theme(legend.position = "none") +
  coord_flip()
```
