---
title: "Significance Testing of Rq3 Groups"
author: "Jeffrey M. Young"
date: "15 May 2020"
output:
  pdf_document:
        latex_engine: xelatex
theme: prettydoc
---

Hello, in this walk-through we'll perform the significance testing on the
financial and automotive datasets in light of the non-normality results from the
`rq1_rq3` walk through. We are assuming you have read that walk-through first.

First, libraries, note that I have silenced the shadowing warnings from R:
```{r message=FALSE}
library(ggplot2) #plotting
library(dplyr)   #dataframe manipulation
library(tidyr)   #dataframe manipulation
library(broom)   #for the tidy function
library(scales)  #for scientific function
```

Let's load the data, we immediately pipe this with dplyr for conveniences like
manipulating the arrow to look nice and making factors. I drop the Name column
because it is large and separated to the other columns. We use the name `Config`
to stand for version variants throughout the scripts.

```{r, data}
finRawFile <- "../../data/fin_rq3_singletons.csv"
autoRawFile <- "../../data/auto_rq3_singletons.csv"

finSingData <- read.csv(file=finRawFile) %>%
  mutate(Algorithm = as.factor(Algorithm), Config = as.factor(Config)) %>%
  mutate(Algorithm = gsub("-->", "\U27f6", Algorithm), data = "Financial") %>%
  group_by(Algorithm, Config) %>%
  mutate(TimeCalc = time -append(0,head(time, -1))) %>% filter(TimeCalc > 0)


autoSingData <- read.csv(file=autoRawFile) %>%
  mutate(Algorithm = as.factor(Algorithm), Config = as.factor(Config)) %>%
  mutate(Algorithm = gsub("-->", "\U27f6", Algorithm), data = "Auto") %>%
  group_by(Algorithm, Config) %>%
  mutate(TimeCalc = time -append(0,head(time, -1))) %>% filter(TimeCalc > 0)
```

## RQ3: Statistically meaningful overhead of the variational solver
In the last walk-through we found that the data was not normally distributed, and
variance was not homogeneous about the sample groups. Thus, we cannot soundly
perform a two-way ANOVA and trust the results. To overcome this we must use a
`non-parametric` test for statistically significant comparison between groups.
We choose to use a Kruskal-Wallis test, as this is commonly accepted as a
`non-parametric` hypothesis test, unfortunately the Kruskal-Wallis test is only
one-way. We begin with `financial` and then perform the test on `auto`. In the
ANOVA analysis we were able to specify the model _with_ an interaction:
`TimeCalc ~ Config * Algorithm`. For the Kruskal-Wallis this is not the case,
rather we must reproduce the analysis manually.

#### Financial
```{r fin.kw.alg}
## performing the test for Algorithm being significant
kruskal.test(TimeCalc ~ Algorithm, finSingData)
```

Algorithms statistically explain variance in the dataset, as expected.


```{r fin.kw.v}
## and for versions
kruskal.test(TimeCalc ~ Config, finSingData)
```

And so do versions, again as expected.


```{r fin.kw.int}
## We must manually construct the interaction
fin.inters <- interaction(finSingData$Algorithm, finSingData$Config)
kruskal.test(TimeCalc ~ fin.inters, finSingData)
```

Interactions are also meaningful. Significant pairs are found with a Wilcox
test:

```{r fin.pairs}
## perform the test
fin.pairs <- pairwise.wilcox.test(finSingData$TimeCalc, fin.inters,
                                  ## choose a bonferroni p-value adjustment for familywise
                                  ## error rate bonferroni is widely accepted, and very
                                  ## conservative, alternatives exist but are left to
                                  ## the interested
                                  p.adj="bonf", exact=FALSE,
                                  paired=FALSE) %>%
  ## cleanup
  tidy %>%
  separate(group1, sep=c(3,4), into = c("AlgLeft", "Dump", "ConfigLeft")) %>%
  separate(group2, sep=c(3,4), into = c("AlgRight", "Dump2", "ConfigRight")) %>%
  select(-Dump, -Dump2) %>%
  ## filter to version variants who were compared with themselves
  filter(ConfigRight == ConfigLeft) %>%
  ## add data column as `Financial` and sort by p-value
  mutate(data = "Financial") %>%
  arrange(p.value)

## view the data frame
fin.pairs
```

We see in the snippet of the `fin.pairs` data frame that all of the
statistically significant comparisons involve `v-->v` (`vsat`), which confirm
our observations from the rq3 violin plot in the `rq1_rq3.Rmd` walk-through. We
now perform the same analysis on `auto`:

#### Automotive

```{r auto.kw.alg}
## Algorithms are significant
kruskal.test(TimeCalc ~ Algorithm, autoSingData)
```

Algorithms are significant, but just by a small margin as shown by the
`p-value`'s proximity to 0.05. Thus, in a more constrictive hypothesis test they
would be considered not significant for the `auto` dataset. This would imply
that `v-->v` is _not_ statistically different from other algorithms for the
`auto` dataset.

```{r auto.kw.v}
## Versions are significant as expected
kruskal.test(TimeCalc ~ Config, autoSingData)
```

```{r auto.kw.int}
## Interaction, also significant as expected
auto.inters <- interaction(autoSingData$Algorithm, autoSingData$Config)
kruskal.test(TimeCalc ~ auto.inters, autoSingData)
```

For convenience we synthesize the `adj.p.value`'s into a dataframe:

```{r auto.pairs}
## Auto pairs which are significant
auto.pairs <- pairwise.wilcox.test(autoSingData$TimeCalc, auto.inters,
                                   p.adj="bonf", method="holm"
                                 , exact=TRUE, paired=FALSE) %>%
  ## cleanup
  tidy %>%
              separate(group1, sep=c(3,4),
                       into = c("AlgLeft", "Dump", "ConfigLeft")) %>%
              separate(group2, sep=c(3,4),
                       into = c("AlgRight", "Dump2", "ConfigRight")) %>%
             select(-Dump, -Dump2) %>%
             ## restricting comparisons to the same versino variant
             filter(ConfigRight == ConfigLeft) %>%
             ## add data column as `Auto`, then sort
             mutate(data = "Auto") %>%
             arrange(p.value)

## observe the dataset
auto.pairs
```

We see that only two comparisons (v-->v, p-->v), and (v-->p, p-->v) for $V_{1}$
are meaningful in the `auto` dataset. This is good evidence the `v-->v` is not
statistically worse for the `auto` dataset. The exact cause behind this result
requires a more robust dataset, ideally, a dataset composed of several product
lines which have a distribution of sharing ratios. Frankly, more data is needed
to assess the signal. Finally, combine the datasets and synthesize a `p-value`
matrix to visualize significant comparisons:

```{r matrix, warning=FALSE, fig.width=8, fig.height=5}
options(scipen = 999)
rq3pvDF <- rbind(auto.pairs, fin.pairs) %>%
  mutate(Significance = case_when(p.value <= 0.05 ~ "Significant",
                                  TRUE ~ "Not Significant"),
         SigColor = paste(AlgLeft,":",Significance,sep=""),
         Version = factor(ConfigLeft, levels = c("V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "V10")))



ggplot(rq3pvDF, aes(x=AlgLeft, y=AlgRight, size=(1-p.value),
                                shape=SigColor, color=Significance)) +
  geom_point() +
  geom_jitter() +
  theme_classic() +
  facet_grid(data ~ Version, scales="free") +
  scale_size_continuous(range=c(2,4)) +
  scale_shape_manual(values = c(6,5,18,2,17)) +
  guides(size=FALSE, shape=FALSE) +
  ggtitle("RQ3: Statistical significance comparison matrix") +
  ylab("Algorithm") +
  theme(panel.grid.major.y = element_line(color = "lightgrey", linetype="dashed"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "bottom")
```

This plot is faceted by `data`, and `Config` or each version. This means that it
is a matrix of subplots, each subplot shows a dot comparing each algorithm on
the x-axis to the algorithm on the y-axis. The points in the plot are filled in,
_and sized_ by the magnitude of the `p-value`, the lower the `p-value`, the
larger the point in the plot. Clearly shown is `v-->v` statistically different
for financial, and similarly the differences previously reported for `auto`,
$V_{1}$. The actual dataset may be observed with a call to `rq3pvDF` as have
been done many times in these walk-throughs for the interested.
