---
title: "Variational Model Diagnostics Significance Testing of Rq3 Groups"
author: "Jeffrey M. Young"
date: "16 May 2020"
output:
  pdf_document:
        latex_engine: xelatex
theme: prettydoc
---

Hello, in this walk-through we'll perform the significance testing on the
financial and automotive datasets in light of the non-normality results from the
`rq1_rq3` walk through. We are assuming you have read that walk-through first.

### Preliminaires
First, libraries, note that I have silenced the shadowing warnings from R:
```{r message=FALSE}
library(ggplot2) #plotting
library(dplyr)   #dataframe manipulation
library(tidyr)   #dataframe manipulation
library(broom)   #for the tidy function
library(scales)  #for scientific function
```

Let's load the data from the `counts` and `diag` files. Most of this should be
pretty familiar.

```{r, data}
## files
finFile        <- "../../data/fin_diag.csv"
autoFile       <- "../../data/auto_diag.csv"
finCountsFile  <- "../../data/fin_counts.csv"
autoCountsFile <- "../../data/auto_counts.csv"

## read the csv and add data column
finCountData  <- read.csv(file=finCountsFile) %>% mutate(data = "Fin")
autoCountData <- read.csv(file=autoCountsFile) %>% mutate(data = "Auto")

## Munge the datasets
autoRatio <- autoCountData %>%
  ## perform the following verbs with respect to variants
  group_by(Variants) %>%
  ## find raw count of Satisfiable models by variant
  count(Satisfiable) %>%
  ## turn the long data into wide data
  pivot_wider(names_from=Satisfiable, values_from=n) %>%
  ## calclulate the unsat ratio
  mutate(UnSatRatio = UnSat / (UnSat + Sat)) %>%
  ## finally replace NAs from calculations with 0s
  replace_na(list(UnSatRatio = 0, UnSat = 0))

## repeat for the financial dataset
finRatio <- finCountData %>%
  group_by(Variants) %>%
  count(Satisfiable) %>%
  pivot_wider(names_from=Satisfiable, values_from=n) %>%
  mutate(UnSatRatio = signif(UnSat / (UnSat + Sat), 3)) %>%
  replace_na(list(UnSatRatio = 0, UnSat = 0))

## Read in the model diagnostics files and calculate the unchanged ratio
## then merge with the previous data frames
finDF <- read.csv(file=finFile) %>%
  mutate(data = "Fin", UnchangedRatio = NumUnchange /NumFeatures ) %>%
  merge(finRatio)

autoDF <- read.csv(file=autoFile) %>%
  mutate(data = "Auto", UnchangedRatio = NumUnchange / NumFeatures) %>%
  merge(autoRatio)
```

And we can view the datasets:

```{r fin.peek}
## for fin
str(finDF)
summary(finDF)
```

and

```{r auto.peek}
## for auto
str(autoDF)
summary(autoDF)
```

To plot we'll merge this into a single data frame and make a line + scatter plot:

```{r plotting, fig.width=7, fig.height=4}
df <- rbind(finDF, autoDF)

## custom breaks by dataset again
breaksRq1 <- function(x) {
  if (max(x) > 16) {
    2^(1:10)
  } else {
    2^(1:4)}
}

## define the x-axis for both scatter and line plots to be variants
ggplot(df, aes(x=Variants)) +
  ## we add `geom` layers to the plots with different y-axis variables
  ## we do not require a second y-axis because we've normalized the y-axis to ratios
  geom_point(aes(y=UnchangedRatio,color="% Features Unchanged"),size=2) +
  geom_point(aes(y=UnSatRatio,color="% Unsatisfiable Models"),size=2) +

  ## for the line plot we change the line type based on what is being plotted
  ## and increase the size
  geom_line(aes(y=UnchangedRatio, color="% Features Unchanged"),
            linetype="dashed", size=1.1) +
  geom_line(aes(y=UnSatRatio, color="% Unsatisfiable Models"),
            linetype="dotdash", size=1.1) +

  ## make x-axis log scale, use our custom breaks function and facet by dataset
  scale_x_log10(breaks=breaksRq1) +
  facet_wrap(. ~ data, scales="free_x") +

  ## niceties such as legend position, and theming
  theme_classic() +
  scale_y_continuous(breaks=seq(0,1,0.1)) +
  theme(legend.position=c(0.80,0.15),
        legend.text=element_text(size=10),
        legend.key.size = unit(.55,'cm')) +
  guides(color=guide_legend("")) +
  ylab("Percent of Total") +
  ggtitle("Ratio of unsatisfiable variants, and constant Features in V-Model")
```
